Scalar Feedback based Joint Time-Frequency Precoder interpolation for MIMO-OFDM Systems

### 2019-10-11 paper/main.tex
Changes by Kumar

\title{Scalar Parametrization for Joint Time-Frequency interpolation in a MIMO-OFDM channel}

@@

To quantize the subcarriers efficiently and exploit the joint
time-frequency interpolation at the transmitter side, we select the
subcarriers for quantization in an alternating fashion as shown in
Fig.~\ref{fig:adpm-fig}. The figure shows the channel evolution of an
$N$ subcarrier MIMO-OFDM system with subcarriers which are going to be
quantized at every $p$-th subcarriers, for even time instances, where,
$p$ is the gap between two quantized subcarrier indices. For odd time
instances, subcarriers at the position $pk+q$ for k = 0,1,...,
$\frac{N-1}{p}$ with q = ${\frac{p}{2}}$. TODO

Therefore, in ADPCM channel tracking scheme, the arrows in the figure show how previous quantized values along with the 1-bit number find the new quantized value. Now with the above discussion, the bit budget for each subcarrier will be equal to number of scalars * 1-bit. On an average, this can be reduced to half if we transmit only $\theta$ values for one-time instance and only $\phi$ values for the next time while keeping the un transmitted values same as the previous one. This was, the average number of bits required to quantize a channel matrix will be $N_{T}(2N_{T} - N_{R})/2$.

One problem associated with the tracking of phases ($\phi$) is that they change abruptly (cite the figure) between $-\pi$ and $\pi$ since $\sin/\cos(\pi) =\sin/\cos(-\pi)$. Therefore we have to wrap around the values by removing those abrupt changes (/cite another figure here).

% We initialize $\hat{x}$ i.e. $\hat{x}_1$ by quantizing the $x_1$ using $\textbf{B}$ bits.

\subsection{Joint Time Frequency Interpolation}
\label{interp}

After obtaining the $\beta$ values(1-bit numbers) at the trasnmitter we will calculate the subcarrier indices at indices $pk$ or $pk+q$. To reproduce the rest of the subcarriers will use future time instances to do the interpolation. As you can see in the figure, each parameter is interpolated using the subcarriers of the next six time instances. We do the interpolation using the spline-cubic interpolation. We could also use other standard techniques by calculating the MMSE at the missing subcarrier locations but the spline-cubic interpolation method gives better results over the other method.

Now here we can see the advantage of using subcarriers in an alternating fashion as given in the previous section. With this method we could exploit both time and frequency correlation because

%Explain prediction and quantization steps here itself ! (succintly)
\section{Simulation and discussion}
\label{section3}

 We have performed simulations to analyse the performance of the proposed quantization and interpolation method for slowly time-varying MIMO OFDM channel. Channel is simulated using the typical Jakes Model for both ITU Vehicular-A channel and ITU Pedestrian Channel. Performance is measured by averaging over 10 channel realizations with 100 simulations each. The normalised doppler frequency $f_{norm} = f_{max}.T_s$ is set to $3.5\times10^{-4}$ with symbol time period $T_d = 5\times10^{-8} s$. For both vehicular and pedestrian channel, $N_T=4$ and $N_R=2$. Therefore, the total number of parameters are $N_{R}(2N_{T} - N_{R}) = 12$ with 5$\theta$’s and 7$\phi$’s. There are 1024 subcarriers in Vehicular channel and 64 subcarriers in the Pedestrian Channel. We take s = 33/9 for Vehicular and pedestrian channel respectively I.e. 32/8 equally spaced subcarriers with indices 33t/9t, t = 0,1, … , 31 and 0,1, .. ,7 respectively.

@@

\subsection{Bilinear Interpolation}
Bilinear interpolation is the equivalent of linear interpolation in two dimensions. On the transmitter side, all the scalars - $\Theta$ and $\Phi$ of the missing subcarriers are interpolatted using the values of the neigbours. These interpolated values and then used to reconstruct the precoding matrices all together.

>>>>

\title{Scalar Feedback based Joint Time-Frequency Precoder
  interpolation for  MIMO-OFDM Systems}

@@

To quantize the subcarriers efficiently and exploit the joint
time-frequency interpolation at the transmitter, we select the
subcarriers for quantization in an alternating fashion as shown in
Fig.~\ref{fig:adpm-fig}. The figure shows the channel evolution of an
$N$ subcarrier MIMO-OFDM system with the subcarriers that are
quantized. Every $p$-th subcarrier starting with $0$ is quantized for
even time instances, where, $p$ is the gap between two quantized
subcarrier indices. For odd time instances, subcarriers at the
position $pk+q$ for $k = 0,1,..., \frac{N-1}{p}$ with $q =
{\frac{p}{2}}$.

For the adaptive channel tracking scheme, the arrows in
Fig.~\ref{fig:adpm-fig} show how previous quantized values along with
the 1-bit enhancement are used to find the new quantized value. Since
we use a single bit adaptation, the amount of feedback required is 1
bit for each scalar parameter that is being tracked. On an average,
this can be reduced by half if we transmit only $\Theta$ values for
one-time instance and only $\Phi$ values for the next time instant in
an alternating fashion (the untransmitted value is assumed to be the
same as that at the previous instant). In this case, the average
number of bits required to quantize a channel matrix will be
$N_{T}(2N_{T} - N_{R})/2$.

One issue that is associated with the tracking the $\phi_i$ parameters
is that they may change abruptly (cite the figure) between $-\pi$ and
$\pi$ due to jumps of $2\pi$. We have avoided this by unwrapping the
phases to facilitate continuous tracking.
\subsection{Joint Time Frequency Interpolation}
\label{interp}
After obtaining the $\beta$ values (using the single bit feedback) at
the transmitter, we can construct the precoder at subcarrier indices
$pk$ or $pk+q$. To construct the rest of the subcarriers, we will use
precoder parameters from both the past and future and time instances
to do the interpolation (using backtracking to ensure causality). In
Fig.~\ref{fig:adpm-fig}, each parameter is interpolated using the
neighbouring subcarriers using bilinear interpolation (equivalent to
linear interpolation in two dimensions). These interpolated values and
then used to reconstruct the precoding matrices all together. Thus,
the approach here is able to exploit the time and frequency
correlation jointly to enhance performance.

One reason why a linear predictive quantization is preferred is
because the underlying parameters can generally be thought to emerge
from an autoregressive (AR) process. For such processes, it is well known
that linear prediction based on past samples is optimal. Even though
the $\Theta$ and $\Phi$ parameters do not manifest as AR processes,
for small changes, a linear approximation works sufficiently
well, as in the description in~\cite{4114278}, although that was
limited to tracking the temporal evolution of the parameters.

\section{Simulation and discussion}
\label{section3}
We present some simulations of the scalar feedback based joint
time-frequency predictive quantization outlined in the previous section. performed simulations to analyse the performance of the proposed quantization and interpolation method for slowly time-varying MIMO OFDM channel. Channel is simulated using the typical Jakes Model for both ITU Vehicular-A channel and ITU Pedestrian Channel. Performance is measured by averaging over 10 channel realizations with 100 simulations each. The normalised doppler frequency $f_{norm} = f_{max}.T_s$ is set to $3.5\times10^{-4}$ with symbol time period $T_d = 5\times10^{-8} s$. For both vehicular and pedestrian channel, $N_T=4$ and $N_R=2$. Therefore, the total number of parameters are $N_{R}(2N_{T} - N_{R}) = 12$ with 5$\theta$’s and 7$\phi$’s. There are 1024 subcarriers in Vehicular channel and 64 subcarriers in the Pedestrian Channel. We take s = 33/9 for Vehicular and pedestrian channel respectively I.e. 32/8 equally spaced subcarriers with indices 33t/9t, t = 0,1, … , 31 and 0,1, .. ,7 respectively.

@@
